\input format.tex

% \tit IOS 17

\chap Chapter 19 Preprocessing directives

\sec 19.3 Macro replacement [cpp.replace]

The identifiers \symbol{\_\_VA\_ARGS\_\_} and \symbol{\_\_VA\_OPT\_\_} shall
occur only in the replacement-list of a function-like macro that uses the
ellipsis notation in the parameters.

A preprocessing directive of the form
\begtt
# define identifier replacement-list new-line
\endtt
defines an {\it object-like} macro... The replacement list is then re-scanned
for more macro names.

A preprocessing directive of the form
\begtt
# define identifier lparen replacement-list_opt ) new-line
\endtt
defines a {\it function-like} macro with parameters.


\sec 19.3.1 Argument substitution [cpp.subst]

A parameter in the replacement list, unless preceded by a \symbol{\#} or
\symbol{\#\#} preprocessing token or followed by a \symbol{\#\#} preprocessing
token, is replaced by the corresponding argument {\bf after} all macros contained
therein have been expanded.

An identifier \symbol{\_\_VA\_ARGS\_\_} that occurs in the replacement list
shall be treated as if it were a parameter, and the variable arguments shall
form the preprocessing tokens used to replace it.

The token sequence \symbol{\_\_VA\_OPT\_\_(content)} shall be treated as if it
were a parameter, and the preprocessing tokens used to replace it are defined as
follows. If the variable arguments consist of no tokens, the replacement
consists of a single placemarker preprocessing token.  Otherwise, the
replacement consists of the results of the expansion of content as the
replacement list of the current function-like macro before rescanning and
further replacement.

% #### 19.3.3 The `#` operator [cpp.stringize]
%
% Each `#` preprocessing token in the replacement list for a function-like macro
% shall we followed by a parameter as next preprocessing token in the replacement
% list.
%
% If, in the replacement list, a parameter is immediately preceded by a `#`
% preprocessing token, both are are replaced by a single character string literal
% preprocessing token that contains the spelling of the preprocessing token
% sequence for the corresponding argument.
%
%
%
% #### 19.3.3 The `##` operator [cpp.concat]
%
% If, in the replacement list of a function-like macro, a parameter is immediately
% preceded or followed by a `##` preprocessing token, the parameter is replaced by
% the corresponding argument's preprocessing token sequence.
%
% > Personal note: This is the reason why to expand `__COUNTER__`, we need two
% > macros, rather than one.
%
%     #include <iostream>
%
%     // The following code does not work.
%     //
%     //    #define CONCAT(x, y) x##y
%     #define CONCAT(x, y) CONCAT_IMPL(x, y)
%     #define CONCAT_IMPL(x, y) x##y
%
%     #define MKSTR(x) # x
%     #define NAME(x) MKSTR(x)
%
%     #define UNIQUE_NAME(x) NAME(CONCAT(x, __COUNTER__))
%     #define PRINT_NEW_NAME
%
%     int main(int argc, char** argv) {
%       std::cout << UNIQUE_NAME(h) << "\n";
%       std::cout << UNIQUE_NAME(h) << "\n";
%       return 0;
%     }
%
% > End of Personal note.
%
% #### 19.3.4 Rescanning and further replacement [cpp.rescan]
%
% After all parameters in the replacement list have been substituted and `#` and
% `##` processing has taken place...Then the resulting preprocessing token
% sequence is rescanned, along with all subsequent preprocessing token of the
% source file.

\vfill
\bye
