\input doc/format
\notespagesize
%
Volume 4
%
\section{Horn Clause and Formula}

A Horn clause is a clause (a disjunction of literals) with at most one positive
literal.  A Horn formula is a propositional formula formed by conjunction of
Horn clauses.

\vskip\baselineskip
\noindent{\bf Definite Horn Formula and The Core}: The definite Horn formula
must satisfy
$$
f(1, 1, ..., 1) = 1.
$$

The core of the definite Horn formula is set the variables which must be true
whenever $f$ is true.

\vskip\baselineskip
\noindent{\bf Algorithrm C}

\noindent The algorithm is as follows

\noindent 1. Put all positive literal in the single variable clause into core.
   They must be true.

\noindent 2. Keep deducing the proposition of non-positive literal, once their values
   are known to be true, i.e., in core. Once all non-positive literals in a
   clause are deduced, its positive literal, if present, must be in core.

\noindent{\bf Indefinite Horn Formula}

Exercise 48 provides the steps to test satisfiability of Horn formula in
general. The idea is quite simple

\noindent 1. Introduce a new variable $\lambda$, and convert all indefinite
causes to definite cause. For example, the following indefinite Horn clause
$$
\bar a \vee \bar b  % \wedge
$$
will be converted as
$$
\bar a \vee \bar b \vee \lambda.
$$
\noindent 2. Apply Algorithrm C to the new definite Horn formula. The original
Horn formula is satisfiable if and only if `lambda` is not in the core of the
new definite Horn formula.

\bye
